hh---
title: "Rosetta"
author: "M Schmettow"
date: "05/10/2022"
output: html_document
---

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(printr)
require(readxl)
require(jpeg)
require(ggimg)
```

## Input and preparation

### Reading meta data

#### Tidyverse

```{r}
library(tidyverse)

read_yeta1_stim <- 
  function(file = "Stimuli/Stimuli.csv", 
           stim_dir = "Stimuli/",
           col_types = cols(File = col_character(),
                            width = col_double(),
                            height = col_double(),
                            hum_like = col_double(),
                            hum_skull = col_logical(),
                            hum_eye = col_logical(),
                            Face = col_character())){
    stimuli <-
      read_csv(file, col_types = col_types) |> 
      mutate(Path = str_c(stim_dir, File, sep = "")) |> 
      rename(Stim = File)
    return(stimuli)    
  }

read_yeta1_aoi <- 
  function(file = "Stimuli/AOI.csv",
           stim_dir = "Stimuli/",
           col_types = cols(AOI = col_character(), Face = col_character(), 
                            x = col_double(), y = col_double(), 
                            w = col_double(), h = col_double())) {
    
    read_csv(file, col_types = col_types) %>% 
      mutate(xmin = x, 
             xmax = x + w,
             ymin = y - h,
             ymax = y) |>  ## reversing the y coordinates) %>%
      rename(x_aoi = x, y_aoi = y, w_aoi = w, h_aoi = h) %>% 
      arrange(Face, AOI)
  }


stims <- read_yeta1_stim() 
aoi <-  read_yeta1_aoi()


```






#### Pandas

```{python eval = FALSE}
from pandas import read_csv
import os

stim_dir = os.path.dirname("Stimuli/")

dtype = {
        'File': 'string',
        'width': 'int16',
        'height': 'int16',
        'hum_like': 'float32',
        'hum_skull': 'boolean',
        'hum_eyes': 'boolean',
        'Face': 'string'
    }
stims = read_csv(file, dtype = dtype)\
          .apply(lambda row: os.path.join(stim_dir, row["File"]),  axis = 1)


stims = read_yeta1_stim()
stims.sample(10)
```

#### Polars

```{python}
import polars as pl
from polars import * #read_csv, String, Int16, Float32, Boolean, col
import os

schema_stim = { 'File': String,
                'width': Int16,
                'height': Int16,
                'hum_like': Float32,
                'hum_skull': Int16,
                'hum_eye': Int16,
                'Face': String}

stim_dir = "Stimuli/"
stims = read_csv(stim_dir + "/Stimuli.csv", schema = schema ).\
          cast({"hum_eye": Boolean, "hum_skull": Boolean}).\
          with_columns((stim_dir + col("File")).alias("Path"))

#stims.sample(10)
```

```{python}
import polars as pl
from polars import * #read_csv, String, Int16, Float32, Boolean, col
import os

schema_aoi={'Face': String,
        'AOI': String,
        'x': Float32,
        'y': Float32,
        'w': Float32,
        'h': Float32}


aois = read_csv(stim_dir + "/AOI.csv", schema = schema_aoi)\
  .with_columns(((col.x).alias("xmin")),
                (col.x + col.w).alias("xmax"),
                (col.y - col.h).alias("ymin"),
                (col.y).alias("ymax"))\
  .rename({"x": "x_aoi",
           "y": "y_aoi",
           "w": "w_aoi",
           "h": "h_aoi"})
```

### Reading raw data

#### Tidyverse

```{r}
library(tidyverse)

read_yeta1_csv <- function(file){
  read_csv(file, 
           col_types = cols(Part = col_character(), 
                            Picture = col_character(), 
                            time = col_double(), 
                            xraw = col_double(), yraw = col_double(), 
                            x = col_double(), y = col_double())) %>% 
    mutate(is_duplicate = x == lag(x) & y == lag(y)) %>% 
    filter(!is_duplicate) %>% 
    mutate(File = Picture, 
           Exp = "UV22") %>% 
    select(Exp, Part, Stim = Picture, File = Picture, time, x, y)
}

csv_files <- dir(path = "Data/",
             pattern = "*.csv",
             recursive = T,
             full.names = T)

add_travel   <- function(data) mutate(data,  travel = sqrt((x - lag(x))^2 + (y - lag(y))^2))
add_duration <- function(data) mutate(data,  duration = lead(time) - time)
is_in_aoi   <-  function(data) mutate(data, is_in_aoi = x > xmin & x < xmax & y > ymin & y < ymax ) 

D_0 <- 
    csv_files %>% 
    map_df(read_yeta1_csv) %>% 
    mutate(Exp = "UV22",
           Obs  = row_number()) %>%
    mutate(Part = as.factor(as.integer(Part) - min(as.integer(Part)) + 1)) %>% ## reducing the Part identifier
    group_by(Part) %>%
    mutate(time = time - min(time)) %>% # time since start experiment
    ungroup() %>%
    left_join(stims, by = "Stim") %>%
    right_join(aoi, by = "Face") |> # this expands the data set 
    add_travel() |> 
    add_duration() |> 
    is_in_aoi() |> 
    filter(is_in_aoi) |> 
    select(Obs, Part, Stim, Face, time, x, y, AOI, hum_like, hum_skull, hum_eye, duration, travel)

colnames(D_0)
dim(D_0)[1]

sample_n(D_0, 10)
```


#### Pandas

```{python}
import pandas as pd

dtypes = 

```


#### Polars

```{python}
from polars import *
from glob import glob
import os

schema_csv={
        'Part': Int32,
        'Picture': String,
        'time': Float64,
        'x':    Float32,
        'y':    Float32,
        'xraw': Float32,
        'yraw': Float32
    }


def read_yeta1_csv(file):
  out = read_csv(file, schema = schema_csv)
  out = out.with_columns(shift_x = col("x").shift(), shift_y = col("y").shift())
  #out = out.rename({"Picture": "Stim"})
  out = out.filter((out["x"]!=out["shift_x"]) & (out["y"]!=out["shift_y"]))
  out = out.drop(["shift_x", "shift_y"])
  return out


def is_in_aoi(data):
  return data.with_columns(col("is_in") = col.x > xmin & col.x < xmax)
    

csv_dir = os.path.join(os.path.curdir, "Data")
csv_files = glob(csv_dir + "/*.csv")    

D_0 = DataFrame(schema = schema_csv)
for f in csv_files:
  D = D_0.vstack(read_yeta1_csv(f))

D = D.rename({"Picture": "Stim"})
D = D.with_columns(Part = (col("Part") - col("Part").min() + 1))
D = D.with_columns(time = (col("time") - col("time").min()).over("Part"))
D = D.with_columns(col("Part").cast(String))


D = D.join(stims, left_on = "Stim", right_on = "File", how = "left")


D = D.join(aois, on = "Face", how = "right", suffix = "_aoi")
D = D.with_columns(is_in_x = (col("x") > col("x_aoi")) & (col("x") < (col("x_aoi") + col("w_aoi"))),
                is_in_y = (col("y") < col("y_aoi")) & (col("y") > (col("y_aoi") - col("h_aoi"))))
D = is_in_aoi(D)                
                
                
D = D.with_columns(is_in = (col("is_in_x") & col("is_in_y")))
D = D.filter(col("is_in"))

#D = D.select(["x", "y", "x_aoi", "y_aoi", "h_aoi", "w_aoi"])
#  .drop(["is_in_x", "is_in_y"])
D


```





### Raw data visualization

```{r}
get_last_part <- function(data){
  last_part <- 
    distinct(data, Part) %>% 
    filter(as.numeric(Part) == max(as.numeric(Part), na.rm = T)) %>% 
    left_join(data, by = "Part")
  return(last_part)
}

```

```{r}
G_0 +
  geom_point(aes(x = x, y = y),
             size = 2,
             col = "red",
             alpha = .2,
             data = get_last_part(D_0),
             inherit.aes = F) +
  facet_wrap(~Stim)
```

```{r}
G_0 +
  geom_point(aes(x = x, y = y,
                 col = Part), # <--
             size = 2,
             data = D_0, # <--
             inherit.aes = F) +
  facet_wrap(~Stim)
```

```{r}
G_0 +
  geom_point(aes(x = x, y = y,
                 col = Part),
             data = get_last_part(D_0),
             size = .1,
             inherit.aes = F) +
  geom_line(aes(x = x , 
                y = y,
                group = Part),
                col = "red",
             inherit.aes = F,
            data = get_last_part(D_0)) +
  facet_wrap(~Stim)
```

### Adding Stimulus meta data

```{r}
D_1 <- 
  D_0 %>% 
  left_join(Stimuli, by = "Stim") %>% 
  select(Obs, Part, Stim, Face, hum_like, Sclera, Skull, time, x, y)

sample_n(D_1, 12)
```

## Deriving measures

-   measuring travel and duration

```{r}
add_travel <- 
  function(data) 
    mutate(data,  travel = sqrt((x - lag(x))^2 + (y - lag(y))^2))

add_duration <- 
  function(data) 
    mutate(data,  duration = lead(time) - time)


D_2 <- 
  D_1 %>% 
  mutate(Sequence = as.factor(str_c(Part, Stim, sep = "_"))) %>% 
  group_by(Sequence) %>% 
  add_duration() %>% 
  add_travel() %>% 
  ungroup() %>% 
  select(Obs, Part, Stim, Face, Sequence, hum_like, 
         Sclera, Skull, time, x, y, travel, duration)
```

```{r}

D_2 %>% 
  ggplot(aes(x = duration)) +
  geom_histogram() +
  facet_wrap(~Stim)


D_2 %>% 
  ggplot(aes(x = travel)) +
  geom_histogram() +
  facet_wrap(~Stim)
```

### Areas of interest

#### Reading AOI

```{r}
read_yeta1_aoi <- 
  function(file = "Stimuli/AOI.csv",
           stim_dir = "Stimuli/",
           col_types = cols(AOI = col_character(), Face = col_character(), 
                            x = col_double(), y = col_double(), 
                            w = col_double(), h = col_double()),
           stim_tab) {
    
    read_csv(file, col_types = col_types) %>% 
      rename(x_aoi = x, y_aoi = y, w_aoi = w, h_aoi = h) %>% 
      right_join(stim_tab, by = "Face") %>% 
      mutate(xmin = x_aoi, 
             xmax = x_aoi + w_aoi,
             ymax = height - y_aoi, ## reversing the y coordinates
             ymin = (height - y_aoi) - h_aoi) %>% 
      arrange(Face, AOI)
  }

```

```{r}
AOI <- read_yeta1_aoi(stim_tab = Stimuli)

head(AOI)
```

#### AOI preview

```{r fig.height = 8, fig.width = 8}
G_1 <- 
  AOI %>% 
  ggplot(aes(xmin = 0, xmax = width, 
             ymin = 0, ymax = height)) +
  facet_wrap(~Face) + # <--
  ggimg::geom_rect_img(aes(img = Path)) +
  geom_rect(aes(xmin = xmin, ymin = ymin, 
                xmax = xmax, ymax = ymax,
                fill = AOI),
            alpha = .2, 
            inherit.aes  = F)

G_1
```

#### AOI Classification

```{r}
D_3 <- 
  D_2 %>% 
  left_join(AOI, by = "Face") %>% 
  mutate(is_in = x > xmin & x < xmax & y > ymin & y < ymax) %>% 
  filter(is_in) %>% 
  select(Obs, AOI) %>% 
  right_join(D_2, by = "Obs") %>% 
  mutate(AOI = if_else(is.na(AOI), "Outside", AOI)) %>% 
  arrange(Part, time)

summary(D_3)
```

```{r}
D_3 %>% 
  group_by(AOI, Sclera, Skull) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  ggplot(aes(y = n, x = AOI, fill = AOI)) +
  facet_grid(Skull~Sclera) +
  geom_col()

```

```{r}
G_0 +
  geom_count(aes(x = x, y = y, 
                 col = AOI),
             alpha = .5,
             inherit.aes  = F,
             data = D_3)
```

### Measuring visits

A *visit* is a closed sequence of eye positions in the same region. The following code uses a combined criterion for setting a new visits:

-   the position falls into a different AOI

-   OR: the travel traveled from the previous position exceeds a certain threshold

```{r}

travel_threshold <- 50

D_4 <-

  D_3 %>%

  group_by(Part, Stim) %>%

  filter(AOI != lag(AOI) | travel > travel_threshold) %>% ## logical OR

  mutate(visit = row_number(),

         duration = lead(time) - time) %>%

  ungroup()

sample_n(D_4, 10)

```

#### Plotting visit paths and duration

```{r fig.width=8, fig.height = 8}

G_3 <-

  G_0 +
  geom_point(aes(x = x, y = y,
                 size = duration), # <--
             color = "white",
             alpha = .2,
             inherit.aes  = F,
             data = D_4)

G_3

```

```{r}

G_4 <-
  G_0 +
  geom_path(aes(x = x, y = y,
                col = Part),
            inherit.aes  = F,
            data = D_4) # <--

G_4

```


## Participant-level analysis

### Frequencies and durations

```{r}

D_6 <-
  D_4 %>%
  group_by(Part, Face, AOI, Sclera, Skull) %>%  # <--
  summarize(n_visits = n(),
            total_dur = sum(duration, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(congruent = (Sclera == Skull))


D_6

```

```{r}

G_6 <-

  D_6 %>%
  ggplot(aes(x = congruent, y = n_visits, fill = AOI)) +
  facet_wrap(~Part) +
  geom_col()

G_6

```


```{r}

G_7 <-
  D_6 %>%
  ggplot(aes(x = AOI, y = total_dur, fill = congruent)) +
  facet_wrap(~Part) +
  geom_col()

G_7

```

```{r}

save(AOI, D_1, D_2, D_3, D_4, D_6, file = "D.Rda")

```

