---
title: "Rosetta"
author: "M Schmettow"
date: "05/10/2022"
output: html_document
---

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(printr)
require(readxl)
require(jpeg)
require(ggimg)
```

## Input and preparation

### Reading meta data

#### Tidyverse

```{r}
library(tidyverse)

read_yeta1_stim <- 
  function(file = "Stimuli/Stimuli.csv", 
           stim_dir = "Stimuli/",
           col_types = cols(File = col_character(),
                            width = col_double(),
                            height = col_double(),
                            hum_like = col_double(),
                            hum_skull = col_logical(),
                            hum_eye = col_logical(),
                            Face = col_character())){
    stimuli <-
      read_csv(file, col_types = col_types) |> 
      mutate(Path = str_c(stim_dir, File, sep = "")) |> 
      rename(Stim = File)
    return(stimuli)    
  }

stims <- 
  read_yeta1_stim()

 
stims |>  sample_n(10)
```

```{r}
read_yeta1_aoi <- 
  function(file = "Stimuli/AOI.csv",
           stim_dir = "Stimuli/",
           col_types = cols(AOI = col_character(), Face = col_character(), 
                            x = col_double(), y = col_double(), 
                            w = col_double(), h = col_double()),
           stim_tab) {
    
    read_csv(file, col_types = col_types) %>% 
      rename(x_aoi = x, y_aoi = y, w_aoi = w, h_aoi = h) %>% 
      right_join(stim_tab, by = "Face") %>% 
      mutate(xmin = x_aoi, 
             xmax = x_aoi + w_aoi,
             ymin = (height - y_aoi) - h_aoi,
             ymax = height - y_aoi) |>  ## reversing the y coordinates) %>% 
      arrange(Face, AOI)
  }

aoi <- read_yeta1_aoi(stim_tab = stims) |> 
  select(Stim, Face, AOI, x_aoi:h_aoi, width , height, hum_like:hum_eye, Path) |> 
  arrange(Stim, AOI)
```






#### Pandas

```{python}
from pandas import read_csv
import os

stim_dir = os.path.dirname("Stimuli/")

dtype = {
        'File': 'string',
        'width': 'int16',
        'height': 'int16',
        'hum_like': 'float32',
        'hum_skull': 'boolean',
        'hum_eyes': 'boolean',
        'Face': 'string'
    }
stims = read_csv(file, dtype = dtype)\
          .apply(lambda row: os.path.join(stim_dir, row["File"]),  axis = 1)


stim = read_yeta1_stim()
stim.sample(10)
```

#### Polars

```{python}
import polars as pl
from polars import * #read_csv, String, Int16, Float32, Boolean, col
import os

schema={
        'File': String,
        'width': Int16,
        'height': Int16,
        'hum_like': Float32,
        'hum_skull': Int16,
        'hum_eye': Int16,
        'Face': String
    }

stim_dir = "Stimuli/"
stims = read_csv(stim_dir + "/Stimuli.csv", schema = schema ).\
          cast({"hum_eye": Boolean, "hum_skull": Boolean}).\
          with_columns((stim_dir + col("File")).alias("Path")).\
          rename

stims.sample(10)
```

```{python}
import polars as pl
from polars import * #read_csv, String, Int16, Float32, Boolean, col
import os

schema={
        'Face': String,
        'AOI': String,
        'x': Float32,
        'y': Float32,
        'w': Float32,
        'h': Float32
    }


aois = read_csv(stim_dir + "/AOI.csv", schema = schema )
```

### Reading raw data

#### Tidyverse

```{r}
library(tidyverse)

read_yeta1_csv <- function(file){
  read_csv(file, 
           col_types = cols(Exp = col_character(), Part = col_character(), 
                            Stim = col_character(), time = col_double(), 
                            x = col_double(), y = col_double(), 
                            x_pro = col_double(), y_pro = col_double())) %>% 
    mutate(is_duplicate = x == lag(x) & y == lag(y)) %>% 
    filter(!is_duplicate) %>% 
    mutate(File = file) %>% 
    select(Exp, Part, Stim, time, x, y, x_pro, y_pro)
}

csv_files <- dir(path = "Data/",
             pattern = "*.csv",
             recursive = T,
             full.names = T)

add_travel <- 
  function(data) 
    mutate(data,  travel = sqrt((x - lag(x))^2 + (y - lag(y))^2))

add_duration <- 
  function(data) 
    mutate(data,  duration = lead(time) - time)


UV22_0 <- 
    csv_files %>% 
    map_df(read_yeta1_csv) %>% 
    mutate(Obs  = row_number()) %>%
    mutate(Part = as.factor(as.integer(Part) - min(as.integer(Part)) + 1)) %>% ## reducing the Part identifier
    group_by(Part) %>%
    mutate(time = time - min(time)) %>% # time since start experiment
    ungroup() %>%
    left_join(stims, by = c("Stim","File")) %>% 
    mutate(y = height - y, # translating from origin at top (pygame) to bottom (ggplot)
           y_pro = 1 - y_pro)


%>%
    select(Exp, Obs, Part, Stim, time, x, y, x_pro, y_pro) |> 
    add_travel() |> 
    add_duration() |> 




```


#### Pandas

```{python}
import pandas as pd

dtypes = 

```


#### Polars

```{python}
from polars import *
from glob import glob

csv_dir = os.path.join(os.path.curdir, "Data")


schema={
        'Exp': String,
        'Part': String,
        'Stim': String,
        'time': Float64,
        'x':    Float32,
        'y':    Float32,
        'x_pro': Float32,
        'y_pro': Float32
    }
    
csv_files = glob(csv_dir + "/*.csv")    

UV22 = DataFrame(schema = schema)
for f in csv_files:
  UV22 = UV22.vstack(read_csv(f, schema = schema))


UV22 = UV22.\
  join(stims, left_on = "Stim", right_on = "File", how = "left").\
  join(aois, on = "Face", how = "left", suffix = "_aoi")


  
```





### Raw data visualization

```{r}
get_last_part <- function(data){
  last_part <- 
    distinct(data, Part) %>% 
    filter(as.numeric(Part) == max(as.numeric(Part), na.rm = T)) %>% 
    left_join(data, by = "Part")
  return(last_part)
}

```

```{r}
G_0 +
  geom_point(aes(x = x, y = y),
             size = 2,
             col = "red",
             alpha = .2,
             data = get_last_part(UV22_0),
             inherit.aes = F) +
  facet_wrap(~Stim)
```

```{r}
G_0 +
  geom_point(aes(x = x, y = y,
                 col = Part), # <--
             size = 2,
             data = UV22_0, # <--
             inherit.aes = F) +
  facet_wrap(~Stim)
```

```{r}
G_0 +
  geom_point(aes(x = x, y = y,
                 col = Part),
             data = get_last_part(UV22_0),
             size = .1,
             inherit.aes = F) +
  geom_line(aes(x = x , 
                y = y,
                group = Part),
                col = "red",
             inherit.aes = F,
            data = get_last_part(UV22_0)) +
  facet_wrap(~Stim)
```

### Adding Stimulus meta data

```{r}
UV22_1 <- 
  UV22_0 %>% 
  left_join(Stimuli, by = "Stim") %>% 
  select(Obs, Part, Stim, Face, hum_like, Sclera, Skull, time, x, y)

sample_n(UV22_1, 12)
```

## Deriving measures

-   measuring travel and duration

```{r}
add_travel <- 
  function(data) 
    mutate(data,  travel = sqrt((x - lag(x))^2 + (y - lag(y))^2))

add_duration <- 
  function(data) 
    mutate(data,  duration = lead(time) - time)


UV22_2 <- 
  UV22_1 %>% 
  mutate(Sequence = as.factor(str_c(Part, Stim, sep = "_"))) %>% 
  group_by(Sequence) %>% 
  add_duration() %>% 
  add_travel() %>% 
  ungroup() %>% 
  select(Obs, Part, Stim, Face, Sequence, hum_like, 
         Sclera, Skull, time, x, y, travel, duration)
```

```{r}

UV22_2 %>% 
  ggplot(aes(x = duration)) +
  geom_histogram() +
  facet_wrap(~Stim)


UV22_2 %>% 
  ggplot(aes(x = travel)) +
  geom_histogram() +
  facet_wrap(~Stim)
```

### Areas of interest

#### Reading AOI

```{r}
read_yeta1_aoi <- 
  function(file = "Stimuli/AOI.csv",
           stim_dir = "Stimuli/",
           col_types = cols(AOI = col_character(), Face = col_character(), 
                            x = col_double(), y = col_double(), 
                            w = col_double(), h = col_double()),
           stim_tab) {
    
    read_csv(file, col_types = col_types) %>% 
      rename(x_aoi = x, y_aoi = y, w_aoi = w, h_aoi = h) %>% 
      right_join(stim_tab, by = "Face") %>% 
      mutate(xmin = x_aoi, 
             xmax = x_aoi + w_aoi,
             ymax = height - y_aoi, ## reversing the y coordinates
             ymin = (height - y_aoi) - h_aoi) %>% 
      arrange(Face, AOI)
  }

```

```{r}
AOI <- read_yeta1_aoi(stim_tab = Stimuli)

head(AOI)
```

#### AOI preview

```{r fig.height = 8, fig.width = 8}
G_1 <- 
  AOI %>% 
  ggplot(aes(xmin = 0, xmax = width, 
             ymin = 0, ymax = height)) +
  facet_wrap(~Face) + # <--
  ggimg::geom_rect_img(aes(img = Path)) +
  geom_rect(aes(xmin = xmin, ymin = ymin, 
                xmax = xmax, ymax = ymax,
                fill = AOI),
            alpha = .2, 
            inherit.aes  = F)

G_1
```

#### AOI Classification

```{r}
UV22_3 <- 
  UV22_2 %>% 
  left_join(AOI, by = "Face") %>% 
  mutate(is_in = x > xmin & x < xmax & y > ymin & y < ymax) %>% 
  filter(is_in) %>% 
  select(Obs, AOI) %>% 
  right_join(UV22_2, by = "Obs") %>% 
  mutate(AOI = if_else(is.na(AOI), "Outside", AOI)) %>% 
  arrange(Part, time)

summary(UV22_3)
```

```{r}
UV22_3 %>% 
  group_by(AOI, Sclera, Skull) %>% 
  summarize(n = n()) %>% 
  ungroup() %>% 
  ggplot(aes(y = n, x = AOI, fill = AOI)) +
  facet_grid(Skull~Sclera) +
  geom_col()

```

```{r}
G_0 +
  geom_count(aes(x = x, y = y, 
                 col = AOI),
             alpha = .5,
             inherit.aes  = F,
             data = UV22_3)
```

### Measuring visits

A *visit* is a closed sequence of eye positions in the same region. The following code uses a combined criterion for setting a new visits:

-   the position falls into a different AOI

-   OR: the travel traveled from the previous position exceeds a certain threshold

```{r}

travel_threshold <- 50

UV22_4 <-

  UV22_3 %>%

  group_by(Part, Stim) %>%

  filter(AOI != lag(AOI) | travel > travel_threshold) %>% ## logical OR

  mutate(visit = row_number(),

         duration = lead(time) - time) %>%

  ungroup()

sample_n(UV22_4, 10)

```

#### Plotting visit paths and duration

```{r fig.width=8, fig.height = 8}

G_3 <-

  G_0 +
  geom_point(aes(x = x, y = y,
                 size = duration), # <--
             color = "white",
             alpha = .2,
             inherit.aes  = F,
             data = UV22_4)

G_3

```

```{r}

G_4 <-
  G_0 +
  geom_path(aes(x = x, y = y,
                col = Part),
            inherit.aes  = F,
            data = UV22_4) # <--

G_4

```


## Participant-level analysis

### Frequencies and durations

```{r}

UV22_6 <-
  UV22_4 %>%
  group_by(Part, Face, AOI, Sclera, Skull) %>%  # <--
  summarize(n_visits = n(),
            total_dur = sum(duration, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(congruent = (Sclera == Skull))


UV22_6

```

```{r}

G_6 <-

  UV22_6 %>%
  ggplot(aes(x = congruent, y = n_visits, fill = AOI)) +
  facet_wrap(~Part) +
  geom_col()

G_6

```


```{r}

G_7 <-
  UV22_6 %>%
  ggplot(aes(x = AOI, y = total_dur, fill = congruent)) +
  facet_wrap(~Part) +
  geom_col()

G_7

```

```{r}

save(AOI, UV22_1, UV22_2, UV22_3, UV22_4, UV22_6, file = "UV22.Rda")

```

