---
title: "YET with R basic workflow"
author: "M Schmettow"
date: "09/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(fig.height = 8)
data_path = "CSV"
```

```{r}
library(tidyverse)
library(printr)

require(readxl)
require(jpeg)
require(ggimg)
```

# Preparation

```{r}
WIDTH = 450
HEIGHT = 450
IMG_DIR = "Stimuli/"
```

## Reading csv

```{r}
# csv_files <- dir(path = data_path,
#                  pattern = "*.csv",
#                  recursive = T,
#                  full.names = T)

csv_files <- "Data/yeta1_UV22MS1663266938.csv"

UV22_0 <- 
  csv_files %>% 
  map_df(~read_csv(.x, 
                   col_types = cols(Experiment = col_character(),
                                    Part = col_double(), # <-- we change this later
#                                    Obs = col_double(),
                                    Stimulus = col_character(),
                                    time = col_double(),
                                    x = col_double(),
                                    y = col_double(),
                                    x_offset = col_double(),
                                    y_offset = col_double()
                   )) %>%
           mutate(File = .x)) %>% 
  mutate(is_duplicate = lag(x) == x & lag(y) == y) %>% ## Yeta1 seems to duplicate measures. This is a bugfix
  filter(!is_duplicate) %>% 
  filter(!str_detect(Stimulus, "dummy")) %>% 
  mutate(Obs  = row_number()) %>% 
  mutate(Part = as.factor(as.integer(Part - min(Part)) + 1)) %>% ## reducing the Part identifier
  group_by(Part) %>% 
  mutate(time = time - min(time)) %>% # time since start experiment
  ungroup() %>% 
  mutate(y = HEIGHT - y, ### reversing the y-axis
         manipulated = stringr::str_detect(Stimulus, "manipulated"),
         Face = stringr::str_extract(Stimulus, "[0-9]+"),
         humlike = as.numeric(Face)) %>% 
  select(Obs, Part, Stimulus, Face, humlike, manipulated, time, x, y)
  
sample_n(UV22_0, 10)

summary(UV22_0, 10)
```


## Reading StimulusInfo and AOI

```{r}
Stimuli <- 
  read_csv(str_c(IMG_DIR, "Stimuli.csv"),
           col_types = cols(File = col_character(),
                            width = col_double(),
                            height = col_double(),
                            humLike = col_double(),
                            humskull = col_logical(),
                            whitesclera = col_logical(),
                            congruency = col_double()
           )) %>% 
  rename(Stimulus = File) %>% 
  mutate(Skull = if_else( humskull, "human", "ape"),
         Sclera = if_else( whitesclera, "human", "ape"),
         congruent = (Sclera == Skull))

UV22_1 <- left_join(UV22_0, Stimuli, 
                    by = "Stimulus") %>% 
  select(Obs, Part, Stimulus, Face, humlike, Sclera, Skull, congruent, time, x, y)

sample_n(UV22_1, 12)
```



```{r}
AOI <- 
  read_csv(str_c(IMG_DIR, "AOI.csv", sep = "")) %>% 
  #right_join(Stimuli, by = "Stimulus") %>% 
  mutate(Face = str_extract(Stimulus, "[0-9]+"),
         Path = str_c(IMG_DIR, Stimulus, sep = ""),
         #Image = map(Path, ~jpeg::readJPEG(.x)),
         xmin = x, 
         xmax = x + w,
         ymax = HEIGHT - y, ## reversing the y coordinates
         ymin = HEIGHT - (y + h)) %>% 
  arrange(Face, AOI) %>% 
  select(Face, AOI, xmin, xmax, ymin, ymax, Path)

head(AOI)
```

## Data preparation

-   measuring distance and duration
-   vertical mirroring off coordinates
-   extracting variables from file names
-   shortening some variables

```{r}
UV22_2 <- 
  UV22_1 %>% 
  mutate(Sequence = as.factor(str_c(Part, Stimulus, sep = "_"))) %>% 
  group_by(Sequence) %>% 
  mutate(distance = sqrt((x - lag(x))^2 + ## Euclidian distance
                           (y - lag(y))^2),
         duration = lead(time) - time) %>% ## duration
  ungroup() %>% 
  select(Obs, Part, Stimulus, Face, Sequence, humlike, Sclera, Skull, congruent, time, x, y, distance, duration) %>% 
  filter(Face != "dummy")

sample_n(UV22_2, 10)

summary(UV22_2)
```

# Visualization

## Grid of Stimuli

We create a re-usable ggplot object G_0 containing a grid of pictures

```{r, fig.height = 8, fig.width = 8}
G_0 <-
  AOI %>% 
  ggplot(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax)) +
  facet_wrap(~Face) +
  ggimg::geom_rect_img(aes(img = Path, xmin = 0, xmax = WIDTH, ymin = 0, ymax = HEIGHT)) +
  xlim(0, WIDTH) +
  ylim(0, HEIGHT)

G_0
```

## Raw measures visualization

```{r}
G_0 +
  geom_point(aes(x = x, y = y, col = Part),
             size = .5,
             alpha = .5,
             inherit.aes  = F,
             data = UV22_2)
```



## AOI visualization

```{r fig.height = 8, fig.width = 8}
G_1 <- 
  G_0 +
  geom_rect(aes(xmin = xmin, ymin = ymin, 
                xmax = xmax, ymax = ymax,
                fill = AOI),
            alpha = .2, 
            inherit.aes  = F)

G_1
```

## AOI Classification

```{r}
UV22_3 <- 
  UV22_2 %>% 
  left_join(AOI, by = "Face") %>% 
  mutate(is_in = x > xmin & x < xmax & y > ymin & y < ymax) %>% 
  filter(is_in) %>% 
  select(Obs, AOI) %>% 
  right_join(UV22_2, by = "Obs") %>% 
  mutate(AOI = if_else(is.na(AOI), "Outside", AOI)) %>% 
  arrange(Part, time)

summary(UV22_3)
```

```{r}
UV22_3 %>% 
  group_by(AOI) %>% 
  summarize(n())

```

```{r}
G_0 +
  geom_point(aes(x = x, y = y, 
                 col = AOI),
             size = .5,
             alpha = .5,
             inherit.aes  = F,
             data = UV22_3)
```

```{r}
G_0 +
  geom_count(aes(x = x, y = y, 
                 col = AOI),
             alpha = .5,
             inherit.aes  = F,
             data = UV22_3)
```

## Measuring visits

A *visit* is a closed sequence of eye positions in the same region. The following code uses a combined criterion for setting a new visits:

-   the position falls into a different AOI
-   OR: the distance traveled from the previous position exceeds a certain threshold


```{r}
distance_threshold <- 50

UV22_4 <- 
  UV22_3 %>%
  group_by(Part, Stimulus) %>% 
  filter(AOI != lag(AOI) | distance > distance_threshold) %>% ## logical OR
  mutate(visit = row_number(),
         duration = lead(time) - time) %>% 
  ungroup()

sample_n(UV22_4, 10)
```

## Plotting visit paths and duration

```{r fig.width=8, fig.height = 8}
G_3 <- 
  G_0 +
  geom_point(aes(x = x, y = y, 
                 shape = Part,
                 size = duration), # <--
             alpha = .5,
             inherit.aes  = F,
             data = UV22_4)

G_3
```

```{r}
G_4 <- 
  G_0 +
  geom_path(aes(x = x, y = y, 
                col = Part), 
            inherit.aes  = F,
            data = UV22_4) # <--
G_4
```

## Population-level AOI frequencies

```{r}
UV22_5 <- 
  UV22_4 %>%
  group_by(AOI, congruent) %>% 
  summarize(n_visits = n(),
            total_dur = sum(duration, na.rm = TRUE)) %>% 
  ungroup()

UV22_5
```

```{r}
G_5 <-
  UV22_5 %>% 
  ggplot(aes(x = AOI, y = n_visits, fill = congruent)) +
  geom_col()

G_5
```

## Frequencies per participant

```{r}
UV22_6 <- 
  UV22_4 %>%
  group_by(Part, Face, AOI, congruent, Sclera, Skull) %>%  # <--
  summarize(n_visits = n(),
            total_dur = sum(duration, na.rm = TRUE)) %>% 
  ungroup()

UV22_6
```


```{r}
G_6 <-
  UV22_6 %>% 
  ggplot(aes(x = congruent, y = n_visits, fill = AOI)) +
  facet_wrap(~Part) +
  geom_col()

G_6
```

## Durations per participant

```{r}
G_7 <-
  UV22_6 %>% 
  ggplot(aes(x = AOI, y = total_dur, fill = manipulated)) +
  facet_wrap(~Part) +
  geom_col()

G_6
```


```{r}
save(AOI, UV22_1, UV22_2, UV22_3, UV22_4, UV22_5, UV22_6, file = "UV22.Rda")
```

# Your analysis


```{r}
load(file = "UV22.Rda")
```


```{r}
library(rstanarm)
options(mc.cores = 4)
library(bayr)
```


## Exploration


```{r}
UV22_4 %>% 
  group_by(AOI, congruent) %>% 
  summarize(mean_dur = mean(duration, na.rm = TRUE),
            sd_dur = sd(duration, na.rm = TRUE))
```


```{r}
M_1 <- 
  UV22_4 %>% 
  stan_glm(duration ~ 1 + AOI + congruent + AOI:congruent, # AOI*congruence
           data = .,
           family = Gamma())

```

```{r}
fixef(M_1)
```

```{r}
M_2 <- 
  UV22_4 %>% 
  stan_glm(duration ~ 0 + AOI:congruent,
           data = .)

```


```{r}
fixef(M_2)
```

```{r}
T_2 <-
  fixef(M_2) %>% 
  mutate(fixef = str_remove_all(fixef, "AOI|congruent")) %>% 
  separate(fixef, into = c("AOI", "congruence")) %>% 
  select(AOI, congruence, center, lower, upper)

T_2
```

```{r}
T_2 %>% 
  ggplot(aes(x = AOI, col = congruence, 
             y = center, ymin = lower, ymax = upper)) +
  geom_point() +
  geom_line(aes(group = congruence))
```

```{r}
M_3 <- stan_glm(n_visits ~ 1 + congruent, 
                data = UV22_5,
                family = neg_binomial_2())
```

```{r}
fixef(M_3, mean.func = exp)
```

```{r}
M_4 <- stan_glm(n_visits ~ 1 + Sclera * Skull, 
                data = UV22_6,
                family = neg_binomial_2())
```


```{r}
fixef(M_4)
```

```{r}
M_5 <- stan_glm(n_visits ~ 0 + Sclera : Skull, 
                data = UV22_6,
                family = neg_binomial_2())
```


```{r}
fixef(M_5, mean.func = exp)
```



